---
title: "Mental Health code"
author: "Sara"
date: "2024-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}
library("shiny")
library("rsconnect")
library("agridat")
library("readxl")
library("ggplot2")
library("dplyr")
library("shinydashboard") 
library("shinythemes")
library("shinyWidgets")
library("ggplot2")
library("DT") 
library("plotly")
```



```{r}
# Read File 
data = read.csv("mental_health_finaldata_1.csv")

```



```{r}
#dimension
dim(data)
```

```{r}
#variables
names(data)
```


```{r}
#Checking Data type of Variable "Age"
class(data$Age)
```

#Checking Variable data type 
#All the variables are Character in the data

```{r}
#Checking Data type of Variable "Gender"
class(data$Gender)
```


```{r}
#Checking Data type of Variable "Occupation"
class(data$Occupation)
```

```{r}
#Checking Data type of Variable "Days_Indoors"
class(data$Days_Indoors)

```

```{r}
#Checking Data type of Variable "Growing_Stress"
class(data$Growing_Stress)


```

```{r}
#Checking Data type of Variable "Quarantine_Frustrations"
class(data$Quarantine_Frustrations)
```


```{r}
#Checking Data type of Variable "Changes_Habits"
class(data$Changes_Habits)
```

```{r}
#Checking Data type of Variable "Mental_Health_History"
class(data$Mental_Health_History)
```


```{r}
#Checking Data type of Variable "Weight_Change"
class(data$Weight_Change)
```



```{r}
#Checking Data type of Variable "Mood_Swings"
class(data$Mood_Swings)
```


```{r}
#Checking Data type of Variable "Coping_Struggles"
class(data$Coping_Struggles)
```



```{r}
#Checking Data type of Variable "Work_Interest"
class(data$Work_Interest)
```



```{r}
#Checking Data type of Variable "Social_Weakness"
class(data$Social_Weakness)
```

#Creating a Dataframe
```{r}
data1<-data.frame(data)
```

#Summary of dataframe
```{r}
summary(data1)
```



#Cleaning the dataframe 
#First we check for the NA values for each variables.
#Than we sum the NA values for the dataframe. There is no NA in the dataframe
#Checked the variables value and rename them appropriately
#A new dataframe is created for variable "Age", 
#For the variable Age we rename them "20-25" = "Early Twenties", "30-Above" = "Thirties and Above", "25-30" = "Late Twenties","16-20" ~ "Late Teens"
#

```{r}
#Checking NA value for variable "Age"
which(is.na(data1$Age))
```


```{r}
#Checking NA value for variable "Gender"
which(is.na(data1$Gender))
```

```{r}
#Checking NA value for variable "Occupation"
which(is.na(data1$Occupation))
```

```{r}
#Checking NA value for variable "Days_Indoors"
which(is.na(data1$Days_Indoors))
```

```{r}
#Checking NA value for variable "Growing_Stress"
which(is.na(data1$Growing_Stress))
```

```{r}
#Checking NA value for variable "Quarantine_Frustrations"
which(is.na(data1$Quarantine_Frustrations))
```

```{r}
#Checking NA value for variable "Changes_Habits"
which(is.na(data1$Changes_Habits))
```

```{r}
#Checking NA value for variable "Mental_Health_History"
which(is.na(data1$Mental_Health_History))
```

```{r}
#Checking NA value for variable "Weight_Change"
which(is.na(data1$Weight_Change))
```

```{r}
#Checking NA value for variable "Mood_Swings"
which(is.na(data1$Mood_Swings))
```

```{r}
#Checking NA value for variable "Coping_Struggles"
which(is.na(data1$Coping_Struggles))
```


```{r}
#Checking NA value for variable "Work_Interest"
which(is.na(data1$Work_Interest))
```



```{r}
#Checking NA value for variable "Social_Weakness"
which(is.na(data1$Social_Weakness))
```


```{r}
#No NA value in the dataframe
sum(is.na(data1))
```



```{r}
#Checking the unique values in the variables Age in the dataframe and renaming them.
#unique(data1$Age)

# Rename values in the Age variable 
# "20-25" = "Early Twenties"
# "30-Above" = "Thirties and Above"
# "25-30" = "Late Twenties"
# "16-20" ~ "Late Teens"
data1$Age[data1$Age == "20-25"] = "20"
data1$Age[data1$Age == "30-Above"] = "30 "
data1$Age[data1$Age == "25-30"] = "25"
data1$Age[data1$Age == "16-20"] = "16"
```




```{r}
#Checking the unique values in the variables Days_Indoors in the dataframe and renaming them.
# Rename observation in the Days_Indoors variable 
# "1-14 days" = "Short-Term"
# "15-30 days" = "Thirties and Above"
# "31-60 days" = "Medium-Term"
# "More than 2 months" ~ "Long-Term"
# "Go out Every day" = "Daily"
unique(data1$Days_Indoors)
data1$Days_Indoors[data1$Days_Indoors == "1-14 days"] = "Short-Term"
data1$Days_Indoors[data1$Days_Indoors == "31-60 days"] = "Medium-Term"
data1$Days_Indoors[data1$Days_Indoors == "Go out Every day"] = "Daily"
data1$Days_Indoors[data1$Days_Indoors == "More than 2 months"] = "Long-Term"
data1$Days_Indoors[data1$Days_Indoors == "15-30 days"] = "Thirties and Above"
```






```{r}
#Checking the unique values in the variables Growing_Stress in the dataframe and renaming them.
# Rename observation in the Growing_Stress 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some stress.
unique(data1$Growing_Stress)
```


```{r}
data1$Growing_Stress[data1$Growing_Stress == "Maybe"] = "Yes"

```

```{r}
#Checking the unique values in the variables  in the dataframe and renaming them.
# Rename observation in the Quarantine_Frustrations 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some Frustrations.
unique(data1$Quarantine_Frustrations)
```


```{r}
data1$Quarantine_Frustrations[data1$Quarantine_Frustrations == "Maybe"] = "Yes"
```



```{r}
#Checking the unique values in the variables  in the dataframe and renaming them.
# Rename observation in the Changes_Habits 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some Changes_Habits.
unique(data1$Changes_Habits)
```


```{r}
data1$Changes_Habits[data1$Changes_Habits == "Maybe"] = "Yes"

```

```{r}
#Checking the unique values in the variables  in the dataframe and renaming them.
# Rename observation in the Mental_Health_History 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some Mental_Health_History.
unique(data1$Mental_Health_History)
```


```{r}
data1$Mental_Health_History[data1$Mental_Health_History == "Maybe"] = "Yes"

```


```{r}
#Checking the unique values in the variables  in the dataframe and renaming them.
# Rename observation in the Weight_Change 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some Weight_Change.
unique(data1$Weight_Change)
```


```{r}
data1$Weight_Change[data1$Weight_Change == "Maybe"] = "Yes"

```




```{r}
#Checking the unique values in the variables  in the dataframe and renaming them.
# Rename observation in the Work_Interest 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some Work_Interest.
unique(data1$Work_Interest)
```


```{r}
data1$Work_Interest[data1$Work_Interest == "Maybe"] = "Yes"
```



```{r}
#Checking the unique values in the variables  in the dataframe and renaming them.
# Rename observation in the Social_Weakness 
# The "Maybe" observations are taken as 'Yes'. Becasuse they are not sure about it, that mean they definitely got some Social_Weakness.
unique(data1$Social_Weakness)
```


```{r}
data1$Social_Weakness[data1$Social_Weakness == "Maybe"] = "Yes"


```

```{r}
data1$Age <- as.numeric(as.character(data1$Age))

```

```{r}
colnames(data1)
```




```{r}
library(shiny)
library(ggplot2)
# Define UI
ui <- fluidPage(
  
  # App title
  titlePanel("Mental Health Data Analysis"),
  
  # Sidebar layout with input and output definitions
  sidebarLayout(
    
    # Sidebar panel for inputs
    sidebarPanel(
      
      selectInput(inputId = "hist_var",
                  label = "Histogram variable:",
                  choices = c("Age"),
                  selected = "Age"),
      
      # Input: Select variable for pie chart
      selectInput(inputId = "pie_var",
                  label = "Pie chart variable:",
                  choices = c("Gender","Days_Indoors", "Occupation", "Growing_Stress", 
                              "Quarantine_Frustrations", "Changes_Habits", 
                              "Mental_Health_History", "Weight_Change", 
                              "Mood_Swings", "Coping_Struggles", 
                              "Work_Interest", "Social_Weakness"),
                  selected = "Gender")
    ),
    
    # Main panel for displaying outputs
    mainPanel(
      plotOutput(outputId = "histogram"),
      plotOutput(outputId = "pie_chart")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  # Generate histogram based on user input
  output$histogram <- renderPlot({
    ggplot(data1, aes_string(x = input$hist_var)) +
      geom_histogram(fill = "lightblue", color = "black", bins = 10) +
      labs(x = input$hist_var, y = "Frequency", title = paste("Histogram of", input$hist_var))
  })
  
  # Generate pie chart based on user input
  output$pie_chart <- renderPlot({
    counts <- table(data1[[input$pie_var]])
    pie(counts, main = paste("Pie Chart of", input$pie_var), col = rainbow(length(counts)))
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```


